---
title: "PredictiveModeling_WaterAccess"
author: "Jeff Grayum"
date: "6/8/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
#install.packages("usemodels")
library(usemodels)
#install.packages("tune")
#install.packages("ranger")
#install.packages("themis")
library(themis)
library(recipes)
library(ranger)
library(ggthemes)
library(knitr)
knitr::opts_chunk$set(cache = TRUE, cache.lazy = FALSE, warning = FALSE, 
                      message = FALSE, echo = TRUE, dpi = 180, 
                      fig.width = 8, fig.height = 5)
theme_set(theme_minimal())
update_geom_defaults("rect", list(fill = "midnightblue", alpha = 0.8))
```
OK, when you get the below error message when using a new package, you must update a *different* package for the new one to function...  Update the package on the third line, in single quotes ('tune').

Error: package or namespace load failed for ‘usemodels’ in loadNamespace(i, c(lib.loc, .libPaths()), versionCheck = vI[[i]]):
 namespace ‘tune’ 0.1.1 is already loaded, but >= 0.1.2 is required
 
 
```{r}
tuesdata <- tidytuesdayR::tt_load('2021-05-04')
water_row <- tuesdata$water
```

We are going to try and build a model to predict whether water is available at a given source.
```{r}
water_row %>%
  count(status_id, sort = TRUE)
#We are going to exclude the Unknown status_id

water_row %>%
  count(water_source, sort = TRUE)
#There is a lot of missing data in this dataset.
```

First, let's make a quick map.  Just to see what's up.
```{r}
water_row %>%
  filter(status_id %in% c("y", "n")) %>%
  ggplot(aes(lon_deg, lat_deg, color = status_id)) +
  geom_point(alpha = 0.2)
#This is a lot of data.  Let's look at only a single country.  Good Ol' Sierra Leone.

water_row %>%
  filter(country_name == "Sierra Leone",
         status_id %in% c("y", "n"),
         lon_deg < -5,
         lat_deg < 15) %>%
  ggplot(aes(lon_deg, lat_deg, color = status_id)) +
  geom_point(alpha = 0.2) +
  labs(x = "Longitude",
       y = "Latitude",
       color = "Status", 
       title = "Water sources in Sierra Leone") +
  coord_fixed() +
  guides(color = guide_legend(override.aes = list(alpha = 1)))
```


```{r}
water <- water_row %>%
  filter(country_name == "Sierra Leone",
         status_id %in% c("y", "n"),
         lon_deg < -5,
         lat_deg < 15) %>%
  mutate(pay = case_when(str_detect(pay, "^No") ~"no",  #Cool str_detect trick
                         str_detect(pay, "^Yes") ~"yes",
                         is.na(pay) ~ pay,
                         TRUE ~ "it's complicated")) %>%
  select(-country_name, -status, -report_date) %>%
  mutate_if(is.character, as.factor) #To model, we must convert our characters to factors.
```

```{r}
water %>%
  ggplot(aes(install_year, y = ..density.., fill = status_id)) +
  geom_histogram(position = "identity", alpha = 0.5) +
  labs(fill = "Water available?")

water %>%
  ggplot(aes(y = pay, fill = status_id)) +
  geom_bar(position = "fill") + #Allows us to look at %'s
  labs(fill = "Water available?")

```

Time to build the model!  Set seed, split, train split, test split.
```{r}
#First, let's set up our spending our data budget (wtf?)

set.seed(123)

water_split <- initial_split(water, strata = status_id) #Strata = what we're trying to predict!

water_train <- training(water_split)

water_test <- testing(water_split)

#Use training data in resampling!
set.seed(234)
water_folds <- vfold_cv(water_train, strata = status_id) 
water_folds

#Above is "spending our data budget".  Nice.
```
Ok, so table below is our resamples (water_folds).  In each resample, we have some data that we're going to use for analysis left(37k), and some that we're going to use for assessment (right, 4.1k)  

A tibble: 10 x 2
   splits             id    
   <list>             <chr> 
 1 <split [37K/4.1K]> Fold01
 2 <split [37K/4.1K]> Fold02
 3 <split [37K/4.1K]> Fold03
 4 <split [37K/4.1K]> Fold04
 5 <split [37K/4.1K]> Fold05
 6 <split [37K/4.1K]> Fold06
 7 <split [37K/4.1K]> Fold07
 8 <split [37K/4.1K]> Fold08
 9 <split [37K/4.1K]> Fold09
10 <split [37K/4.1K]> Fold10


We will fit this using a fairly simple model, the Ranger Random Forest Model.  No tuning required!  We have enough trees. " . " means "use everything".  So, we're explaining status_id by all other variables in our dataset.

```{r}
usemodels::use_ranger(status_id ~ ., data = water_train)

#Below, we're just copying and pasting the output from above! Then, making some small changes. step_unknow assings a missing value to unknown

#Below is our recipe for pre-processing and feature engineering.
ranger_recipe <- 
  recipe(formula = status_id ~ ., data = water_train) %>%
  update_role(row_id, new_role = "id") %>%
  step_unknown(all_nominal_predictors()) %>%
  step_other(all_nominal_predictors(), threshold = 0.03) %>%
  step_impute_linear(install_year) %>%
  step_downsample(status_id)
  

#Random forest model specification
ranger_spec <- 
  rand_forest(trees = 1000) %>% 
  set_mode("classification") %>% 
  set_engine("ranger") 

#Added both the feature engineering recipe and model specification to a workflow.
ranger_workflow <- 
  workflow() %>% 
  add_recipe(ranger_recipe) %>% 
  add_model(ranger_spec) 

doParallel::registerDoParallel()
set.seed(74403)
#Now we fit our workflow to our folds.
ranger_tune <- 
  fit_resamples(ranger_workflow, 
                         resamples = water_folds,
                         control = control_resamples(save_pred = TRUE))

```

